\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-05-10 slides/slides.tex
[ slides ] last changes.

\setbeamerfont{framesubtitle}{size=\normalfont\tiny}
\setbeamercolor{framesubtitle}{fg=white, bg=plum!80!black}

@@

\title[Proof Reconstruction in Classical Propositional Logic]
  {\textbf{Proof Reconstruction in Classical Propositional Logic}}
\subtitle{(Work in Progress)}
\date{Agda Implementors’ Meeting XXV\\
May 9-15th 2017}
\author[Jonathan Prieto-Cubides]{Jonathan Prieto-Cubides\\
(Joint work with Andr\'es Sicard-Ram\'irez)}
\institute{
Master in Applied Mathematics\\
Universidad EAFIT\\
Medell\'in, Colombia}
% --------------------------------------------------------------------------

@@

\begin{frame}[plain]
\titlepage
  \begin{tikzpicture}[overlay, remember picture]
   \tikzset{shift={(current page.center)}}
    \node[xshift=-4.5cm,yshift=-3.5cm] (eafit)
      {\includegraphics[width=0.2\textwidth]{figures/eafit}};
    \node[xshift=4.5cm,yshift=-3cm] (chalmers)
      {\includegraphics[width=0.18\textwidth]{figures/chalmers}};
  \end{tikzpicture}
\end{frame}

@@

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{\url{http://www.gilith.com/software/metis}}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{\url{http://github.com/jonaprieto/athena}}
  };
}
\only<6->{
\node[left = 0.6cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{\url{http://github.com/agda/agda}}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<9->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

@@

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] version in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}

@@

\begin{itemize}
  \item $p ⊢ p$
\begin{tptp}
fof(a, axiom, p).
fof(goal, conjecture, p).
\end{tptp}

@@

\begin{frame}[label=metis]{\texttt{Metis} Theorem Prover
  \footnote{\url{http://www.gilith.com/software/metis/}}}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \hyperlink{metis-inference-rules}{\textit{six rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs detailed proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=metis-rules]{Inference Rules of \texttt{Metis}}
\hyperlink{tstp-example}{TSTP} derivations by \texttt{Metis} exhibit these inferences
\footnote{Inference rules found in proofs of Propositional Logic theorems}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Rule} & \textbf{Purpose} \\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF\\
\texttt{clausify} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction\\
\texttt{negate} & applies negation to the formula\\
\texttt{resolve} & applies theorems of resolution\\
\texttt{simplify} & applies over a list of formula to simplify them\\
\texttt{strip} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

@@

A \texttt{TSTP} derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a \hyperlink{tstp-dag}{\textbf{D}irected \textbf{A}cyclic \textbf{G}raph} where
  \begin{itemize}
    \item[\texttt{leaf}] is a formula from the \hyperlink{tptp-syntax}{\texttt{TPTP}} input
    \item[\texttt{node}] is a formula inferred from parent formula
    \item[\texttt{root}] the final derived formula
  \end{itemize}
  \item Is a list of annotated formulas with the form
  \end{itemize}

@@

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents)
\end{tptp}
}
\end{center}
\gobackreconstruction
\end{frame}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} Prover for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG Example}

\vfill
By refutation, we proved $p ⊢ p$:
\begin{center}
\begin{prooftree}
\AxiomC{$p$}
\RightLabel{\texttt{\tiny assume}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\tiny strip}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\tiny negate}}
\UnaryInfC{$¬ p$}
\AxiomC{$p$}
\RightLabel{\texttt{\tiny canonicalize}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\tiny simplify}}
\BinaryInfC{$⊥$}
\RightLabel{\texttt{\tiny canonicalize}}
\UnaryInfC{$⊥$}

@@

\begin{frame}[fragile, label=athena]{\texttt{Athena}
  \footnote{\url{http://github.com/jonaprieto/athena}}}

@@

Is a \texttt{Haskell} program
that translates proofs given by \texttt{Metis} Prover
in \texttt{TSTP} format to \texttt{Agda} code\\

\begin{itemize}
    \item Parsing of \texttt{TSTP} language\footnote{\label{parsing}\url{https://github.com/agomezl/tstp2agda}}\textsuperscript{,}\footnote{\url{https://github.com/jonaprieto/tstp2agda}}
    \item Creation\textsuperscript{\ref{parsing}} and analysis of \hyperlink{tstp-dag}{\textbf{DAG}} derivations
    \item Analysis of inference rules used in the \texttt{TSTP} derivation
    \item \texttt{Agda} code generation
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Library} & \textbf{Purpose} \\
\hline
\texttt{\hyperlink{agda-prop}{\texttt{Agda-Prop}}}& axioms and theorems of Classical Propositional Logic\\
\texttt{\hyperlink{agda-metis}{\texttt{Agda-Metis}}} & versions of the inference rules used by \texttt{Metis}\\
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}

@@

\begin{frame}[fragile, label=agda-prop-2]{\texttt{Agda-Prop} Library
  \footnote{\url{https://github.com/jonaprieto/agda-prop}}}
\begin{itemize}
  \item Intuitionistic Propositional Logic + PEM ($Γ ⊢ ϕ ∨ ¬ ϕ$)
  \item A data type for formulas
  \vspace*{1mm}

\begin{minted}{cagda}
data Prop : Set where
  Var  : Fin n → Prop           -- Variables.
  ⊤    : Prop                   -- Top (truth).
  ⊥    : Prop                   -- Bottom (falsum).
  _∧_  : (φ ψ : Prop) → Prop    -- Conjunction.
  _∨_  : (φ ψ : Prop) → Prop    -- Disjunction.
  _⇒_  : (φ ψ : Prop) → Prop    -- Implication.
  _⇔_  : (φ ψ : Prop) → Prop    -- Biimplication.
  ¬_   : (φ : Prop)   → Prop    -- Negation.
\end{minted}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-prop]{\texttt{Agda-Prop} Library
  \footnote{\url{https://github.com/jonaprieto/agda-prop}}}
\begin{itemize}
  \item A data type for theorems
  \vspace*{1mm}
\begin{minted}{cagda}
data _⊢_ : (Γ : Ctxt)(φ : Prop) → Set
\end{minted}

@@

\begin{frame}[fragile, label=agda-metis]{\texttt{Agda-Metis} Library
  \footnote{\url{https://github.com/jonaprieto/agda-metis}}}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.7}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|lll|}
\hline
\textbf{Rule} & \textbf{Purpose} &\textbf{Theorem}\\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF &\texttt{atp-canonicalize}\\
\texttt{clausify} & performs clausification &\texttt{atp-clausify}\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction &\hyperlink{atp-conjunct}{\texttt{atp-conjunct}}\\
\texttt{negate} & applies negation to the formula &\texttt{atp-negate}\\
\texttt{resolve} & applies theorems of resolution &\texttt{atp-resolve}\\
\texttt{simplify} & applies over a list of formula to simplify them &\texttt{atp-simplify}\\
\texttt{strip} & splits a formula into subgoals &\texttt{atp-strip}\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:agda-metis-table}
\end{table}
\vfill
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{\texttt{Agda-Metis}: Conjunct Inference
  \footnote{\url{https://github.com/jonaprieto/agda-metis}}}
\begin{itemize}
\item Definition
  \begin{equation*}
  \text{conjunct}(ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ \phi_n, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

\item Function\footnote{\label{note-conjunct}Excerpt from the \texttt{Agda-Metis}
library available in  \texttt{ATP.Metis.Rules.Conjunct} module}:
\vspace*{1mm}
\begin{minted}{cagda}
conjunct : Prop → Prop → Prop
conjunct (φ ∧ ψ) ω with ⌊ eq φ ω ⌋ | ⌊ eq ψ ω ⌋
... | true  | _     = φ
... | false | true  = ψ
... | false | false = conjunct φ ω
conjunct φ ω        = φ
\end{minted}
\item Theorem\textsuperscript{\ref{note-conjunct}}
\vspace*{1mm}
\begin{minted}{cagda}
atp-conjunct
  : ∀ {Γ} {φ}
  → (ω :Prop)
  → Γ ⊢ φ
  → Γ ⊢ conjunct φ ω
\end{minted}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile, label=verified-example]{$p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\begin{minted}{cagda}
p, q, a, goal, subgoal₀, subgoal₁ : Prop

-- Axiom.

a = (p ∧ q)

-- Premise.

Γ : Ctxt
Γ = [ a ]

-- Conjecture.

goal = (q ∧ p)

-- Subgoals.

subgoal₀ = q
subgoal₁ = (q ⇒ p)
\end{minted}
\end{frame}

@@

\begin{frame}[fragile, label=verified-example-4]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\begin{minted}{cagda}
-- Premise.

Γ = [ a ]

-- Conjecture.

goal = (q ∧ p)

-- Subgoals.

subgoal₀ = q
subgoal₁ = (q ⇒ p)

-- Proof

proof₀ : Γ ⊢ subgoal₀
proof₁ : Γ ⊢ subgoal₁

proof : Γ ⊢ goal
proof =
  ⇒-elim
    atp-splitGoal            -- q ∧ (q ⇒ p) ⇒ p
    (∧-intro proof₀ proof₁)
\end{minted}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug\footnote{\url{https://github.com/gilith/metis/issues/2}} in the Printing of the Proof}
  {\texttt{Metis}' v2.3 (release 20161108)}
% {}
\begin{tptp}
$ cat problem.tptp
fof(goal, conjecture,
  ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
\end{tptp}
\begin{tptp}
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {\texttt{Metis}' v2.3 (release 20161108) - Issue\footnote{\url{https://github.com/gilith/metis/issues/2}}}

@@

\RightLabel{\scriptsize{simplify}}
\TrinaryInfC{$\bot$}
\end{prooftree}
\pause
The bug was caused by the conversion of \texttt{Xor} sets to \texttt{Iff} lists.
After reporting this, Joe fixed the printing of \texttt{canonicalize} inference rule
\[ φ := ¬ p ∧ (¬ q ⇔ ¬ r) ∧ (¬ p ⇔ (¬ q ⇔ {\color{red} r}))\]
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=hammer]{Related Work}
\begin{block}{SledgeHammer} \citep{Paulson2007}
\begin{itemize}
\item \texttt{Isabelle}/\texttt{HOL} mature tool
\item \texttt{Metis} ported within \texttt{Isabelle}
\item Reconstruct proofs of well-known ATPs: \texttt{EProver}, \texttt{Vampire},
among others using \texttt{SystemOnTPTP} server
\end{itemize}
\end{block}

\begin{block}{Integrating \texttt{Waldmeister} into \texttt{Agda}} \citep{Foster2011}
\begin{itemize}
  \item Framework for a integration between \texttt{Agda} and ATPs
  \begin{itemize}
    \item Equational Logic
    \item Reflection Layers
  \end{itemize}
  \item Source code is not available\footnote{\url{http://simon-foster.staff.shef.ac.uk/agdaatp}}
  \end{itemize}
\end{block}
\end{frame}

@@

\only<2->{
  \node[right=1cm of agdafile] (eagda){
  \includegraphics[scale=0.8]{figures/eagda}
  \footnote{\url{https://github.com/asr/eagda}}
  }
};

@@

\only<3->{
  \node[below=0.4cm of agdai] (apia){
  \includegraphics[scale=0.8]{figures/apia}
  \footnote{\url{https://github.com/asr/apia}}}
  };

@@

\only<5->{ \node[below = 0.8cm of apia] (atp)
  {\includegraphics[scale=0.8]{figures/atp}
  \footnote{\url{http://github.com/jonaprieto/online-atps}}}
  };

>>>>

\setbeamerfont{framesubtitle}{size=\normalfont\scriptsize}
\setbeamercolor{framesubtitle}{fg=white, bg=plum!80!black}

@@

\title[Proof Reconstruction in Classical Propositional Logic]
  {\textbf{Proof Reconstruction in Classical Propositional Logic}}
\subtitle{(work in progress)}
\date{Agda Implementors’ Meeting XXV\\
Teknikparken, Chalmers, Gothenburg, Sweden\\
May 11 2017\\
}
\author[Jonathan Prieto-Cubides]{Jonathan Prieto-Cubides\\
(joint work with Andr\'es Sicard-Ram\'irez)}
\institute{
Master in Applied Mathematics\\
Universidad EAFIT\\
Medell\'in, Colombia}
% --------------------------------------------------------------------------

@@

\begin{frame}[plain]
\titlepage
  \begin{tikzpicture}[overlay, remember picture]
   \tikzset{shift={(current page.center)}}
    \node[xshift=-4.7cm,yshift=-3.2cm] (eafit)
      {\includegraphics[width=0.2\textwidth]{figures/eafit}};
    \node[xshift=4.7cm,yshift=-3cm] (chalmers)
      {\includegraphics[width=0.17\textwidth]{figures/chalmers}};
  \end{tikzpicture}
\end{frame}

@@

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{\url{http://www.gilith.com/software/metis}.}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{\url{http://github.com/jonaprieto/athena}.}
  };
}
\only<6->{
\node[left = 0.6cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{\url{http://github.com/agda/agda}.}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<9->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

@@

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}.}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] formula in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}

@@

\begin{itemize}
  \item $p ⊢ p$
\begin{tptp}
fof(myaxiom, axiom, p).
fof(goal, conjecture, p).
\end{tptp}

@@

\begin{frame}[label=metis]{\texttt{Metis} Theorem Prover
  \footnote{\url{http://www.gilith.com/software/metis/}.}}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \hyperlink{metis-inference-rules}{\textit{six rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs \textit{detailed} proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=metis-rules]{Inference Rules of \texttt{Metis}}
\hyperlink{tstp-example}{\texttt{TSTP}} derivations by \texttt{Metis} exhibit these inferences\footnote{Inference rules found in proofs of Propositional Logic theorems.}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Rule} & \textbf{Purpose} \\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF\\
\texttt{clausify} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction\\
\texttt{negate} & applies negation to the formula\\
\texttt{resolve} & applies theorems of resolution\\
\texttt{simplify} & applies over a list of formula to simplify them\\
\texttt{strip} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

@@

A \texttt{TSTP} derivation\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}.}
\begin{itemize}
  \item Is a \hyperlink{tstp-dag}{\textbf{D}irected \textbf{A}cyclic \textbf{G}raph} where
  \begin{itemize}
    \item[\texttt{leaf}] is a formula from the \hyperlink{tptp-syntax}{\texttt{TPTP}} input
    \item[\texttt{node}] is a formula inferred from parent formula
    \item[\texttt{root}] the final derived formula
  \end{itemize}
  \item Is a list of annotated formulas with the form
  \end{itemize}

@@

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents).
\end{tptp}
}
\end{center}
\gobackreconstruction
\end{frame}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG Example}
\vfill
By refutation, we proved $p ⊢ p$:
\begin{center}
\begin{prooftree}
\AxiomC{$p$}
\RightLabel{\texttt{\scriptsize assume}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize strip}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize negate}}
\UnaryInfC{$¬ p$}
\AxiomC{$p$}
\RightLabel{\texttt{\scriptsize canonicalize}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize simplify}}
\BinaryInfC{$⊥$}
\RightLabel{\texttt{\scriptsize canonicalize}}
\UnaryInfC{$⊥$}

@@

\begin{frame}[fragile, label=athena]{\texttt{Athena}
  \footnote{\url{http://github.com/jonaprieto/athena}.}}

@@

Is a \texttt{Haskell} program
that translates proofs given by \texttt{Metis}
in \texttt{TSTP} format to \texttt{Agda} code\\

\begin{itemize}
    \item Parsing of \texttt{TSTP} language\footnote{\label{parsing}\url{https://github.com/agomezl/tstp2agda}.}\textsuperscript{,}\footnote{\url{https://github.com/jonaprieto/tstp2agda}.}
    \item Creation\textsuperscript{\ref{parsing}} and analysis of \hyperlink{tstp-dag}{\textbf{DAG}} derivations
    \item Analysis of inference rules used in the \texttt{TSTP} derivation
    \item \texttt{Agda} code generation
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Library} & \textbf{Purpose} \\
\hline
\texttt{\hyperlink{agda-prop}{\texttt{Agda-Prop}}}& axioms and theorems of Classical Propositional Logic\\
\texttt{\hyperlink{agda-metis}{\texttt{Agda-Metis}}} & versions of the inference rules used by \texttt{Metis}\\
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}

@@

\begin{frame}[fragile, label=agda-prop-2]{\texttt{Agda-Prop} Library
  \footnote{\url{https://github.com/jonaprieto/agda-prop}.}}
\begin{itemize}
  \item Intuitionistic Propositional Logic + PEM ($Γ ⊢ ϕ ∨ ¬ ϕ$)
  \item A data type for formulas
  \vspace*{1mm}

\begin{minted}{cagda}
data Prop : Set where
  Var  : Fin n → Prop
  ⊤    : Prop
  ⊥    : Prop
  _∧_  : (φ ψ : Prop) → Prop
  _∨_  : (φ ψ : Prop) → Prop
  _⇒_  : (φ ψ : Prop) → Prop
  _⇔_  : (φ ψ : Prop) → Prop
  ¬_   : (φ : Prop)   → Prop
\end{minted}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-prop]{\texttt{Agda-Prop} Library
  \footnote{\url{https://github.com/jonaprieto/agda-prop}.}}
\begin{itemize}
  \item A data type for theorems
  \vspace*{1mm}
\begin{minted}{cagda}
data _⊢_ : (Γ : Ctxt)(φ : Prop) → Set
\end{minted}

@@

\begin{frame}[fragile, label=agda-metis]{\texttt{Agda-Metis} Library
  \footnote{\url{https://github.com/jonaprieto/agda-metis}.}}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.7}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|lll|}
\hline
\textbf{Rule} & \textbf{Purpose} &\textbf{Theorem}\\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF &\texttt{atp-canonicalize}\\
\texttt{clausify} & performs clausification &\texttt{atp-clausify}\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction &\hyperlink{atp-conjunct}{\texttt{atp-conjunct}}\\
\texttt{negate} & applies negation to the formula &\texttt{atp-negate}\\
\texttt{resolve} & applies theorems of resolution &\texttt{atp-resolve}\\
\texttt{simplify} & applies over a list of formula to simplify them &\texttt{atp-simplify}\\
\texttt{strip} & splits a formula into subgoals &\texttt{atp-strip}\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:agda-metis-table}
\end{table}
\vfill
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{\texttt{Agda-Metis}: Conjunct Inference
  \footnote{\url{https://github.com/jonaprieto/agda-metis}.}}
\begin{itemize}
\item Definition
  \begin{equation*}
  \text{conjunct}(ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ \phi_n, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

\item Function
\vspace*{1mm}
\begin{minted}{cagda}
conjunct : Prop → Prop → Prop
conjunct (φ ∧ ψ) ω with ⌊ eq φ ω ⌋ | ⌊ eq ψ ω ⌋
... | true  | _     = φ
... | false | true  = ψ
... | false | false = conjunct φ ω
conjunct φ ω        = φ
\end{minted}
\item Theorem
\vspace*{1mm}
\begin{minted}{cagda}
atp-conjunct
  : ∀ {Γ} {φ}
  → (ω : Prop)
  → Γ ⊢ φ
  → Γ ⊢ conjunct φ ω
\end{minted}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile, label=verified-example]{$p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
\begin{minted}{cagda}
p, q, a, goal, subgoal₀, subgoal₁ : Prop

-- Axiom.
a = (p ∧ q)

-- Premise.
Γ : Ctxt
Γ = [ a ]

-- Conjecture.
goal = (q ∧ p)

-- Subgoals.
subgoal₀ = q
subgoal₁ = (q ⇒ p)
\end{minted}
\vfill
\end{frame}

@@

\begin{frame}[fragile, label=verified-example-4]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\vfill
\begin{minted}{cagda}
-- Premise.
Γ = [ a ]

-- Conjecture.
goal = (q ∧ p)

-- Subgoals.
subgoal₀ = q
subgoal₁ = (q ⇒ p)

-- Proof
proof₀ : Γ ⊢ subgoal₀
proof₁ : Γ ⊢ subgoal₁

proof : Γ ⊢ goal
proof =
  ⇒-elim
    atp-splitGoal            -- q ∧ (q ⇒ p) ⇒ p
    (∧-intro proof₀ proof₁)
\end{minted}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug\footnote{\url{https://github.com/gilith/metis/issues/2}.} in the Printing of the Proof}
  {\texttt{Metis} v2.3 (release 20161108)}
% {}
\begin{tptp}
$ cat problem.tptp
fof(goal, conjecture,
  ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
\end{tptp}
\begin{tptp}
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=failure-example-2]{Bug in the Printing of the Proof}
  {\texttt{Metis} v2.3 (release 20161108)}

@@

\RightLabel{\scriptsize{simplify}}
\TrinaryInfC{$\bot$}
\end{prooftree}
\pause
The bug was caused by the conversion of \texttt{Xor} sets to \texttt{Iff} lists.
After reporting this, Hurd fixed the printing of \texttt{canonicalize} inference rule
\[ φ := ¬ p ∧ (¬ q ⇔ ¬ r) ∧ (¬ p ⇔ (¬ q ⇔ {\color{red} r}))\]
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=hammer]{Related Work}
\begin{block}{SledgeHammer} \citep{Paulson2007}
\begin{itemize}
\item \texttt{Isabelle}/\texttt{HOL} mature tool
\item \texttt{Metis} ported within \texttt{Isabelle/HOL}
\item Reconstruct proofs of well-known ATPs: \texttt{EProver}, \texttt{Vampire},
among others using \texttt{SystemOnTPTP} server
\end{itemize}
\end{block}

\begin{block}{Integrating \texttt{Waldmeister} into \texttt{Agda}} \citep{Foster2011}
\begin{itemize}
  \item Framework for a integration between \texttt{Agda} and ATPs
  \begin{itemize}
    \item Equational Logic
    \item Reflection Layers
  \end{itemize}
  \item Source code is not available\footnote{\url{http://simon-foster.staff.shef.ac.uk/agdaatp}.}
  \end{itemize}
\end{block}
\end{frame}

@@

\only<2->{
  \node[right=1cm of agdafile] (eagda){
  \includegraphics[scale=0.8]{figures/eagda}
  \footnote{\url{https://github.com/asr/eagda}.}
  }
};

@@

\only<3->{
  \node[below=0.4cm of agdai] (apia){
  \includegraphics[scale=0.8]{figures/apia}
  \footnote{\url{https://github.com/asr/apia}.}}
  };

@@

\only<5->{ \node[below = 0.8cm of apia] (atp)
  {\includegraphics[scale=0.8]{figures/atp}
  \footnote{\url{http://github.com/jonaprieto/online-atps}.}}
  };
